### 2018-07-16 manuscript.rmd
incorporating trees and diet data

0 -- PHYLOGENY
0.0 ML ribosomal marker phylogeny constrained by transcriptome tree with BS values at the nodes.
1 -- DESCRIPTIVE MORPHOLOGY AND EVOLUTIONARY HISTORY
1.0 Table showing for each species the mean, number of individuals measured, and standard error of every character, ordered by cnidomic index / tentillum size, and/or colored by clade.
1.05 Plate with all the tentillum diversity.
1.1 Table showing each character’s (illustrating the character per row)  the phylogenetic signal K with the p-value obtained by testing against 1000 simulated permutations, and the number of species included.
1.2 Table showing the best AICc supported model generating the data for each character and its AICc score. Model adequacy.
1.3 Table/figures showing the reconstructed ancestral values at the internal nodes of the phylogeny of characters with a supported phylogenetic model of evolution.
1.4 Figure showing major categorical character shifts, gains, and losses on a simplified phylogeny.
2 -- COMPARATIVE ANALYSES
2.0 Confusion matrix table showing PIC_correlations and trait correlation R2 and p-values for each pair of continuous characters. Confusion matrix for quantitative and categorical characters GLS. pGLS estimation of regression parameters in significantly correlated pairs.
2.1 Table showing the model parameters for each significant regression.
2.3 Verbally PCA and phylogenetic PCA/PICVarA_PCA of the characters. Phylomorphospace superimposed to the tip coordinates in PC1--PC2.
2.4 Table showing normalized rates of evolution for each character BAMM.
3 -- RELATIONSHIPS WITH DIET / FUNCTION / Hypothesis testing
3.0 Phylogeny with the tips mapped with the literature dietary associations, and internal nodes showing SIMMAP reconstructed states.
3.1 Table showing quantitative and categorical characters association effect size with each prey type in the diet.
3.2 Table showing the dAICc of OU model fitting versus white_noise and BM using reconstructed diets as regimes in a 16spp tree. Evaluate model adequacy of OU fits.
3.3 Interpret qualitatively the kinematic implications (high speed video) of the different regions of the morphospace.

>>>>

0 -- PHYLOGENY
0.0 ML ribosomal marker phylogeny constrained by transcriptome tree with BS values at the nodes.
1 -- DESCRIPTIVE MORPHOLOGY AND EVOLUTIONARY HISTORY

```{r load_trees, include=F}
#Load tree
rRNA18S = read.nexus('bipartitions_root.boot10')
transcriptome = read.tree('ultimatephylo2018.tre')
#plot.phylo(consensus)
#is.rooted(consensus)
#nodelabels()
rRNA18S = reroot(rRNA18S, 76, 0.001)
rRNA18S$tip.label = str_replace_all(rRNA18S$tip.label,'_',' ')
rRNA18S$tip.label[rRNA18S$tip.label=="Cordagalma cordiforme"] = "Cordagalma ordinatum"
rRNA18S$tip.label[rRNA18S$tip.label=="Cordagalma sp "] = "Cordagalma ordinatum"
rRNA18S$tip.label[rRNA18S$tip.label=="Clausophyes ovata"] = "Kephyes ovata"
rRNA18S$tip.label[rRNA18S$tip.label=="Erenna sp"] = "Erenna sirena"
rRNA18S$tip.label[rRNA18S$tip.label=="Apolemia sp"] = "Apolemia lanosa"
rRNA18S$tip.label[rRNA18S$tip.label=="Apolemia sp "] = "Apolemia rubriversa"
rRNA18S$tip.label[rRNA18S$tip.label=="Sphaeronectes gracilis"] = "Sphaeronectes koellikeri"
rRNA18S$tip.label[rRNA18S$tip.label=="Prayidae D27SS7"] = "Desmophyes haematogaster"
rRNA18S$tip.label[rRNA18S$tip.label=="Prayidae D27D2"] = "Craseoa lathetica"
rRNA18S = drop.tip(rRNA18S, 41:42) #drop outgroups
rRNA18S = drop.tip(rRNA18S, c(16,31,23,24)) #drop Cordagalma, Clausophyid, Nectopyramis, Nectadamas
#plot(rRNA18S)

transcriptome$tip.label = str_replace_all(transcriptome$tip.label,'_',' ')
transcriptome$tip.label[transcriptome$tip.label=="Cordagalma sp"] = "Cordagalma ordinatum"
transcriptome$tip.label[transcriptome$tip.label=="Prayidae D27SS7"] = "Desmophyes haematogaster"
transcriptome$tip.label[transcriptome$tip.label=="Prayidae D27D2"] = "Craseoa lathetica"
transcriptome = reroot(transcriptome, 22, 0.001)
transcriptome = drop.tip(transcriptome, 19:26) #drop outgroups
#plot(transcriptome)

multitree = list(rRNA18S,transcriptome)
class(multitree)="multiPhylo"
STree = superTree(multitree, rooted = T, method = "MRP", start = transcriptome)
consensus = STree[[2]]
consensus = drop.tip(consensus, "Physonect sp ")
consensus = drop.tip(consensus, "Erenna sirena")
consensus$tip.label[consensus$tip.label=="Physophora gilmeri"] = "Physophora h" ### ARTIFIIAL SUBSTITUTION ###
consensus = drop.tip(consensus, "Physophora hydrostatica")
consensus$tip.label[consensus$tip.label=="Physophora h"] = "Physophora hydrostatica"
consensus$tip.label[consensus$tip.label=="Vogtia pentacantha"] = "Vogtia serrata"
consensus = drop.tip(consensus, "Rosacea flaccida")
#plot(consensus)

#Prune quant matrix to tree species
matrix = castnumbers[which(!is.na(castnumbers$Species)),]
matrix_logs = castlogs[which(!is.na(castlogs$Species)),]
sharedspp = matrix$Species[which(matrix$Species %in% consensus$tip.label)]
sharedmatrix = matrix[which(matrix$Species %in% sharedspp),]
sharedlogs = matrix_logs[which(matrix$Species %in% sharedspp),]
sharedmeans = castmeans[which(castmeans$Species %in% sharedspp),]
sharedmean_logs = castmean_logs[which(castmean_logs$Species %in% sharedspp),]
sharedvars = castvariances[which(castvariances$Species %in% sharedspp),]
sharedvar_logs = castvariance_logs[which(castvariance_logs$Species %in% sharedspp),]
#sharedmeans = solidSPPmeans[which(solidSPPmeans$Species %in% sharedspp),]
#sharedvars = solidSPPvars[which(solidSPPvars$Species %in% sharedspp),]

#Prune categorical matrix to tree species
cat_rowNAs = apply(categorical[,-1], 1, function(x) sum(is.na(x)))
catmatrix = categorical[which(cat_rowNAs<1),] %>% .[,-1]
sharedspp_cat = rownames(catmatrix)[which(rownames(catmatrix) %in% consensus$tip.label)]
sharedcategorical = catmatrix[which(rownames(catmatrix) %in% sharedspp_cat),]

#Prune tree to quant matrix species
nodatatipnames = consensus$tip.label[which(!(consensus$tip.label %in% sharedmatrix$Species))]
nodatatips = c(1:length(consensus$tip.label))[which(consensus$tip.label %in% nodatatipnames)]
prunedtree = drop.tip(consensus, nodatatips)
prunedtree$tip.label
plot(prunedtree)
ultram = chronos(prunedtree)
plot(ultram)

##Make the pseudoindividuals tree
newtree = ultram
sharedmatrix_plied = sharedmatrix
sharedmatrix_plied$slide_id = paste(sharedmatrix$slide_id, sharedmatrix$Species, sep = "_")
for (i in 1:nrow(sharedmatrix_plied)) {
  pos = which(newtree$tip.label == sharedmatrix_plied$Species[i])
  print(pos)
  newtree = bind.tip(newtree, sharedmatrix_plied$slide_id[i], where = pos, edge.length = 0.1)
  print(newtree$tip.label)
  newtree$tip.label = str_replace_all(newtree$tip.label,'_',' ')
}
inditree = drop.tip(newtree, which(!grepl(".[0-9].*", newtree$tip.label)))
zerotipTree = inditree
inditree$edge.length[inditree$edge.length==0] <- 0.0000000001
shavedtree = inditree
inditree = chronos(inditree)
plot(inditree)
sharedmatrix_plied$slide_id = str_replace_all(sharedmatrix_plied$slide_id,'_',' ')
sharedmatrix = sharedmatrix_plied[match(sharedmatrix_plied$slide_id, inditree$tip.label),]
rownames(sharedmatrix) <- 1:nrow(sharedmatrix)

#Prune tree to categorical matrix species
nodatatipnames_cat = consensus$tip.label[which(!(consensus$tip.label %in% sharedspp_cat))]
nodatatips_cat = c(1:length(consensus$tip.label))[which(consensus$tip.label %in% nodatatipnames_cat)]
cat_tree = drop.tip(consensus, nodatatips_cat)
cat_tree$tip.label
plot(cat_tree)
ultram_cat = chronos(cat_tree)
plot(ultram_cat)
sharedcategorical = sharedcategorical[match(ultram_cat$tip.label, rownames(sharedcategorical)),]
cprunedmatrix = sharedcategorical[which(rownames(sharedcategorical)%in%rownames(sharedmatrix)),] %>% .[which(sapply(.,function(x) length(unique(x)))>1)]
sharedbinary = sharedcategorical
sharedbinary$Haploneme.type = as.character(sharedbinary$Haploneme.type)
sharedbinary[sharedbinary=="Isorhizas"] = 0
sharedbinary[sharedbinary=="Anisorhizas"] = 1
sharedbinary$Haploneme.type = as.numeric(sharedbinary$Haploneme.type)
sharedbinary$Heteroneme.type = as.character(sharedbinary$Heteroneme.type)
sharedbinary[sharedbinary=="Stenotele"] = 0
sharedbinary[sharedbinary=="Microbasic mastigophore"] = 1
sharedbinary[sharedbinary=="Eurytele"] = NA
sharedbinary$Heteroneme.type = as.numeric(sharedbinary$Heteroneme.type)

#purely numerics
Q_sharedmatrix = sharedmatrix[,c(-1,-2,-3)]
rownames(sharedmeans) = sharedmeans$Species
Q_sharedmeans = sharedmeans[,-1]
rownames(sharedvars) = sharedvars$Species
Q_sharedvars = sharedvars[,-1]
Q_sharedmean_logs = sharedmean_logs[,-1]
rownames(Q_sharedmean_logs) = sharedmean_logs$Species

```

1.0 Table showing for each species the mean, number of individuals measured, and standard error of every character, ordered by cnidomic index / tentillum size, and/or colored by clade.

```{r export_table_morph, include=FALSE}
morphdata = cbind(castN, castmeans[,-1], castSE[,-1])
morphdata = monrphdata[,c(1:3,33,4,34,5,35,6,36,7,37,8,38,9,39,10,40,11,41,12,42,13,43,14,44,15,45,16,46,17,47,18,48,19,49,20,50,21,51,22,52,23,53,24,54,25,55,26,56,27,57,28,58,29,59,30,60,31,61,32,62)]
names(morphdata) = str_replace_all(names(morphdata),".1","_SE")
write.csv(morphdata, "characterdata.csv")
```

1.05 Plate with all the tentillum diversity.
1.1 Table showing each character’s (illustrating the character per row)  the phylogenetic signal K with the p-value obtained by testing against 1000 simulated permutations, and the number of species included.
1.2 Table showing the best AICc supported model generating the data for each character and its AICc score. Model adequacy.
1.3 Table/figures showing the reconstructed ancestral values at the internal nodes of the phylogeny of characters with a supported phylogenetic model of evolution.
1.4 Figure showing major categorical character shifts, gains, and losses on a simplified phylogeny.
2 -- COMPARATIVE ANALYSES
2.0 Confusion matrix table showing PIC_correlations and trait correlation R2 and p-values for each pair of continuous characters. Confusion matrix for quantitative and categorical characters GLS. pGLS estimation of regression parameters in significantly correlated pairs.
2.1 Table showing the model parameters for each significant regression.
2.3 Verbally PCA and phylogenetic PCA/PICVarA_PCA of the characters. Phylomorphospace superimposed to the tip coordinates in PC1--PC2.
2.4 Table showing normalized rates of evolution for each character BAMM.
3 -- RELATIONSHIPS WITH DIET / FUNCTION / Hypothesis testing

```{r diet_data, include=F}
#Retrieve diet info from literature BINARY
GC = read.csv("Cmerged.csv", header = T, sep=',')[,c(2,4,5)] %>% .[which(grepl("diet",.$character) & .$state==1),]
GC$character = factor(GC$character, levels=unique(GC$character))
GC = GC[which(GC$species %in% sharedmeans$Species),]
GC = split(GC,GC$character)
nrowGC = purrr::map(GC,nrow) %>% as.numeric()
GC = GC[which(nrowGC>2)]
GC = purrr::map(GC,unique)
diet= matrix(ncol=length(GC),nrow=length(unique(sharedmeans$Species))) %>% as.data.frame()
names(diet) = names(GC)
rownames(diet) = unique(sharedmeans$Species)
for(E in GC){
  print(E$species)
  for(S in E$species){
    diet[which(rownames(diet) == S),as.character(unique(E$character))] = 1
  }
}
diet[is.na(diet)] <- 0
diet = diet[which(rowSums(diet)>0),which(colSums(diet)<nrow(diet))]
#diet = rbind(diet, c(0,1,0,0,0,0,0,0))
#rownames(diet)[12] = 'Erenna richardi'
dprunedmatrix = sharedmeans[which(sharedmeans$Species%in%rownames(diet)),]
dprunedmatrix_logs = sharedmean_logs[which(sharedmean_logs$Species%in%rownames(diet)),]
dprunedTree = drop.tip(ultram, which(!(ultram$tip.label %in% rownames(diet))))
plot(dprunedTree)

#Retrieve diet info from literature PROPORTION RELATIVE
quantDiet = read.csv("Qmerged.csv", header = T, sep=',')[,c(2,4,6)]
PP = quantDiet[which(grepl("Relative",quantDiet$character)),]
PP$character = factor(PP$character, levels=unique(PP$character))
PP = PP[which(PP$species %in% sharedmeans$Species),]
PP = dcast(PP, species~character, value.var = "state", mean)
PP[is.na(PP)]<-0
rownames(PP) = PP$species
PP=PP[,-1]
#PP = rbind(PP, c(0,100,0,0,0,0,0,0))
#rownames(PP)[10] = "Erenna richardi"
Pprunedmatrix=sharedmeans[which(sharedmeans$Species%in%rownames(PP)),]
Pprunedmatrix_logs = sharedmean_logs[which(sharedmean_logs$Species%in%rownames(PP)),]
Pprunedtree = drop.tip(ultram, which(!(ultram$tip.label %in% rownames(PP))))
plot(Pprunedtree)

#Retrieve copepod prey length from literature
preylength = quantDiet[which(grepl("prey",quantDiet$character)),] %>% .[,c(1,3)]
names(preylength) = c("Species","Copepod prey length (mm)")
PL_pruned_matrix = sharedmeans[which(sharedmeans$Species %in% preylength$Species),] %>% data.frame(preylength[which(preylength$Species %in% sharedmeans$Species),2])
names(PL_pruned_matrix)[ncol(PL_pruned_matrix)] <- "Copepod prey length (mm)"
```

3.0 Phylogeny with the tips mapped with the literature dietary associations, and internal nodes showing SIMMAP reconstructed states.
3.1 Table showing quantitative and categorical characters association effect size with each prey type in the diet.
3.2 Table showing the dAICc of OU model fitting versus white_noise and BM using reconstructed diets as regimes in a 16spp tree. Evaluate model adequacy of OU fits.
3.3 Interpret qualitatively the kinematic implications (high speed video) of the different regions of the morphospace.
